Boundary terms in the action for Causal Sets

### 2014-10-13 paper_v7.tex
some edits to paper

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$. With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$. If there does exist a $p\in M^+\setminus\Sigma$ and a $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then as $q\in J^-(\Sigma)$ we find that $p\in J^-(\Sigma)$, and hence it is in the set $M^-=J^-(\Sigma)$ by definition. The intersection, $M^+\cap M^-=\Sigma$, is now only true if $p\in\Sigma$ which it does not, and hence there cannot exist such a $p$.  We also have that $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in natural units given by
\be\label{eq:GHYBT_in_continuum}
{S}^{(d)}_{GHY}\left[M^\pm\right]= \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\ee
where $l_p$ is the rationalised Planck length for a general dimension, $l_p=(8\pi G)^{\frac{1}{d-2}}$ (where $G$ is the gravitational constant for a general dimension) and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$. (We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$). Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
\be\label{eq:normal_deriv_boundary}
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$ (corresponding to the rate of change of the surface volume \emph{backwards} in time, as $n^{\mu}$ is past-pointing).

@@

We can also take the mean and limit of $\textbf{S}^{ (d)}_{GHY}$ to find
\begin{gather}\label{eq:boundary_final_proof}
\begin{aligned}
\lim_{\rho\rightarrow\infty} & \left\langle \textbf{S}^{ (d)}_{GHY}\right\rangle = \lim_{\rho\rightarrow\infty} \frac{\rho^{\frac{2}{d}-1}}{l_p^{d-2}} c_{d}\left (\sum_m p_m \left\langle\textbf{F}_m\right\rangle  + \sum_n q_n \left\langle\textbf{P}_n\right\rangle\right)
\\
= & \lim_{\rho\rightarrow\infty}\Bigg[\frac{\rho^{\frac{1}{d}}}{l_p^{d-1}}\frac{c_d}{b_d}\left(\sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}
\\
& +\frac{1}{l_p^{d-2}}\left(\sum_m\frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O(\rho^{-\frac{1}{d}})\Bigg]
\\
& = \frac{1}{l_p^{d-2}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}K
\end{aligned}
\end{gather}
Where the last line is true only if the vectors $\vec{p}$ and $\vec{q}$ satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}. The proof of~\eqref{eq:mainconjecture} is now complete.

>>>>

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$, of finite volume, which admits a \emph{spacelike partitioning} surface, $\Sigma$, defined as a spacelike closed compact surface in $M$, where the causal past and future sets, $M^\pm=J^\pm (\Sigma)$, of the surface then satisfy $M^+\cap M^-=\Sigma$. With this condition it follows that no points in $M^+\setminus\Sigma$ precede any points in $M^-\setminus\Sigma$. If there does exist a $p\in M^+\setminus\Sigma$ and a $q\in M^-\setminus\Sigma$ such that $p\preceq q$, then as $q\in J^-(\Sigma)$ we find that $p\in J^-(\Sigma)$, and hence it is in the set $M^-=J^-(\Sigma)$ by definition. The intersection, $M^+\cap M^-=\Sigma$, is now only true if $p\in\Sigma$ which it does not, and hence there cannot exist such a $p$.  We also have that $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in natural units given by
\be\label{eq:GHYBT_in_continuum}
{S}^{(d)}_{GHY}\left[M^\pm\right]= \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\ee
where $l_p$ is the rationalised Planck length for a general dimension, $l_p=(8\pi G)^{\frac{1}{d-2}}$ (where $G$ is the gravitational constant for a general dimension) and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal covector $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$. (We work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$). Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$:
\be\label{eq:normal_deriv_boundary}
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$ (corresponding to the rate of change of the surface volume \emph{backwards} in time, as $n^{\mu}$ is past-pointing).

@@

We can also take the mean and limit of $\textbf{S}^{ (d)}_{GHY}$ and subsitute in~\eqref{eq:nmax_nmin_final} to find
\begin{gather}\label{eq:boundary_final_proof}
\begin{aligned}
\lim_{\rho\rightarrow\infty} & \left\langle \textbf{S}^{ (d)}_{GHY}\right\rangle = \lim_{\rho\rightarrow\infty} \frac{\rho^{\frac{2}{d}-1}}{l_p^{d-2}} c_{d}\left (\sum_m p_m \left\langle\textbf{F}_m\right\rangle  + \sum_n q_n \left\langle\textbf{P}_n\right\rangle\right)
\\
= & \lim_{\rho\rightarrow\infty}\Bigg[\frac{\rho^{\frac{1}{d}}}{l_p^{d-1}}\frac{c_d}{b_d}\left(\sum_m p_m \frac{\Gamma\left (\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left (\frac{1}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}
\\
+ &\frac{1}{l_p^{d-2}}\left(\sum_m\frac{\Gamma\left (\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left (\frac{2}{d}+n \right)}{n!} \right) \int_{\Sigma}d^{d-1}x\: \sqrt{h}K + O(\rho^{-\frac{1}{d}})\Bigg]
\\
= & \frac{1}{l_p^{d-2}}\int_{\Sigma}d^{d-1}x\: \sqrt{h}K
\end{aligned}
\end{gather}
Where the last line is true only if the vectors $\vec{p}$ and $\vec{q}$ satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}. The proof of~\eqref{eq:mainconjecture} is now complete.
